Modeling memory and taxis to analyze migrations

### 2021-01-16 code/IndicesEquations.Rmd
Changed X.mid to X

where $\sigma_{max}=\frac{1}{\sqrt{12}}.$ The larger the value of social cohesion indicates that the migratory population remained close with one another. 
```{r}
 getSD <- function(f, world){
    EX <- sum( world$X.mid * f) * world$dx
    EX2 <- sum( world$X.mid^2 * f) * world$dx
    sd  = sqrt(EX2 - EX^2)
  }

  computeCohesiveness <- function(P){
    SDs <- apply(P, 1, getSD, world = world)
    SD.max <- (1/sqrt(12))*max(world$X.mid)
    SC <- 1 - SDs/SD.max
    c(SC.mean = mean(SC), SC.sd = sd(SC))
  }

>>>>

where $\sigma_{max}=\frac{1}{\sqrt{12}}.$ The larger the value of social cohesion indicates that the migratory population remained close with one another. 
```{r}
 getSD <- function(f, world){
    EX <- sum( world$X * f) * world$dx
    EX2 <- sum( world$X^2 * f) * world$dx
    sd  = sqrt(EX2 - EX^2)
  }

  computeCohesiveness <- function(P){
    SDs <- apply(P, 1, getSD, world = world)
    SD.max <- (1/sqrt(12))*max(world$X)
    SC <- 1 - SDs/SD.max
    c(SC.mean = mean(SC), SC.sd = sd(SC))
  }

### 2021-01-16 code/StrategicVsTactical_PartI_v2.Rmd
Fixed pulsed resource code and added Bhattacharyya coefficient code
Changed X.mid to X

Out model is an advection diffusion with two advective terms corresponding to a direct response to the environment and to memory.  Importantly, the environment is *approximately periodic*, i.e. one in which the distribution of the resource at a given time is approximately equal to that resource distribution at a fixed previous time period, i.e. $h(t) \approx h(t - \tau)$. This models annual variations, and the period corresponds to one year.  The *tactical* process is straightforward taxis up a resource gradient, while the *collective memory* corresponds to a tendency of the organisms to repeat the behavior of the population in previous years.  Thus:

@@

  X.edge <- seq(X.min, X.max, dx)
  X.mid <- (X.edge[-1] + X.edge[-length(X.edge)])/2
  pop <- outer(Time, X.mid,
                     function(t,x) 
                       dnorm(x, mean = X.max/2 + (peak.max-peak.min) / 2 * sin( 2*(t * pi)/max(Time)),sd = sd))
  pop <- apply(pop, 1, function(x) x/sum(x)/dx) %>% t
  list(pop = pop, X.mid = X.mid, time = Time, dx = dx, tau = tau)
}

@@

fr1 <- dnorm(world$X.mid, mean = 80, sd = 10) %>% {./sum(.)/world$dx}
fr2 <- dnorm(world$X.mid, mean = 20, sd = 10) %>% {./sum(.)/world$dx}
fixed.Resource <- matrix((fr1 + fr2)/2, byrow = TRUE, 
                         nrow = nrow(world$pop), 
                         ncol = ncol(world$pop))

@@

```{r fixedResource, fig.height = 3, fig.width = 4}
with(world, 
  image(X.mid, 1:tau, resource, main = "resource", ylab = "x")
)
```

@@

world.R1 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X.mid, c(t.peak = 25, t.sd = 9, x.peak = 80, x.sd = 9)))

world.R2 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X.mid, c(t.peak = 25, t.sd = 9, x.peak = 80, x.sd = 5)))

world.R3 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X.mid, c(t.peak = 25, t.sd = 5, x.peak = 80, x.sd = 9)))

world.R4 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X.mid, c(t.peak = 25, t.sd = 5, x.peak = 80, x.sd = 5)))

par(mfrow = c(2,2), mar = c(3,3,3,1), mgp = c(2,.25,0), las = 1, tck = 0.01)
with(world.R1, {     
  image(1:tau,X.mid,resource, main = "Broad, long peak")
  contour(1:tau,X.mid,resource, add = TRUE)
  })
with(world.R2, {     
  image(1:tau,X.mid,resource, main = "Narrow, long peak")
  contour(1:tau,X.mid,resource, add = TRUE)
  })
with(world.R3, {     
  image(1:tau,X.mid,resource, main = "Broad, short peak")
  contour(1:tau,X.mid,resource, add = TRUE)
  })
with(world.R4, {     
  image(1:tau,X.mid,resource, main = "Narrow, short peak")
  contour(1:tau,X.mid,resource, add = TRUE)
  })

>>>>

Our model is an advection diffusion with two advective terms corresponding to a direct response to the environment and to memory.  Importantly, the environment is *approximately periodic*, i.e. one in which the distribution of the resource at a given time is approximately equal to that resource distribution at a fixed previous time period, i.e. $h(t) \approx h(t - \tau)$. This models annual variations, and the period corresponds to one year.  The *tactical* process is straightforward taxis up a resource gradient, while the *collective memory* corresponds to a tendency of the organisms to repeat the behavior of the population in previous years.  Thus:

@@

  X.edge <- seq(X.min, X.max, dx)
  X <- (X.edge[-1] + X.edge[-length(X.edge)])/2
  pop <- outer(Time, X,
                     function(t,x) 
                       dnorm(x, mean = X.max/2 + (peak.max-peak.min) / 2 * sin( 2*(t * pi)/max(Time)),sd = sd))
  pop <- apply(pop, 1, function(x) x/sum(x)/dx) %>% t
  list(pop = pop, X = X, time = Time, dx = dx, tau = tau)
}

@@

fr1 <- dnorm(world$X, mean = 80, sd = 10) %>% {./sum(.)/world$dx}
fr2 <- dnorm(world$X, mean = 20, sd = 10) %>% {./sum(.)/world$dx}
fixed.Resource <- matrix((fr1 + fr2)/2, byrow = TRUE, 
                         nrow = nrow(world$pop), 
                         ncol = ncol(world$pop))

@@

```{r fixedResource, fig.height = 3, fig.width = 4}
with(world, 
  image(X, 1:tau, resource, main = "resource", ylab = "x")
)
```

@@

world.R1 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X, c(t.peak = 25, t.sd = 9, x.peak = 80, x.sd = 9)))

world.R2 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X, c(t.peak = 25, t.sd = 9, x.peak = 80, x.sd = 5)))

world.R3 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X, c(t.peak = 25, t.sd = 5, x.peak = 80, x.sd = 9)))

world.R4 <- world %>% 
  list_modify(resource = getPulsedResource(1:world$tau, world$X, c(t.peak = 25, t.sd = 5, x.peak = 80, x.sd = 5)))

par(mfrow = c(2,2), mar = c(3,3,3,1), mgp = c(2,.25,0), las = 1, tck = 0.01)
with(world.R1, {     
  image(1:tau,X,resource, main = "Broad, long peak")
  contour(1:tau,X,resource, add = TRUE)
  })
with(world.R2, {     
  image(1:tau,X,resource, main = "Narrow, long peak")
  contour(1:tau,X,resource, add = TRUE)
  })
with(world.R3, {     
  image(1:tau,X,resource, main = "Broad, short peak")
  contour(1:tau,X,resource, add = TRUE)
  })
with(world.R4, {     
  image(1:tau,X,resource, main = "Narrow, short peak")
  contour(1:tau,X,resource, add = TRUE)
  })
